
\documentclass[fleqn,addpoints]{exam}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xfrac}
\usepackage{float}
\usepackage{amsmath}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}

% \printanswers

\ifprintanswers \usepackage{2in1, lscape} \fi

\title{Math 115 Chapter One Review Problems}
\date{November 23, 2010}
\author{}

\begin{document}

\maketitle

\begin{questions}

\section{Inequalities}

\question $x^2 - 9x \geq 5$
\begin{solution}
\end{solution}

\question $\dfrac{3x+2}{x-1} < 1$
% \vspace{2cm}

\question $|3x-1| > 0$
% \vspace{2cm}

\section{Equations of Circles}
\question Find the equation for a circle with center $(1, -1)$ and radius 10.
\question What is the center and radius of the circle with equation: $x^2+2x+y^2-4y=-1$

\begin{solution}
\[
  (x+1)^2 + (y-2)^2 = 4
\]

\end{solution}
% \vspace{2cm}

\section{Quadratic Functions}

\subsection{Standard Form}
Put each function in standard form and find the vertex.

\question $f(x) = -x^2+4x-3$
\begin{solution}
  $f(x) = -(x-2)^2 + 1$
\end{solution}

\question $f(x) = 6x^2 + 12x - 5$
\begin{solution}
  $f(x) = 6(x+1)^2 - 11$
\end{solution}

\subsection{Graphing}
Sketch the graph of each function.
\question $f(x) = x^2-4x+5$
\question $f(x) = - \dfrac{1}{2} x^2 - x - 2$

\section{Functions}

\subsection{Domain and Range}
Find the domain and range of each function.

\question $f(x) = \dfrac{2}{3x-7}$
% \vspace{2 cm}

\question $f(x) = \sqrt{x^2-3x+2}$
% \vspace{2 cm}

\question $f(x) = \dfrac{1}{\sqrt{x^2+x-4}}$
% \vspace{2 cm}

\subsection{Inverse Functions}

\question
Which of these functions is one-to-one?
\begin{parts}
\part $f(x) = x^3 - 1$
% \vspace{1 cm}
\part $f(x) = 2x + 17$
% \vspace{1 cm}
\part $f(x) = |x-1|$
% \vspace{1 cm}

\end{parts}

\question Find $f^{-1}(x)$
\begin{parts}
  \part $f(x) = 5x+7$
  \part $f(x) = \dfrac{1}{x^2}$, $x > 0$
  % \vspace{2 cm}
  \part $f(x) = \dfrac{2x-1}{x+1}$
% \vspace{3 cm}
\end{parts}

\question
Suppose Pagliacci Pizza charges a base price of \$16 for a large pizza plus \$2 for each topping.  Thus, if you order a
large pizza with $x$ toppings, the price of your pizza is given by the function: $f(x) = 16 + 2x$.  Find $f^{-1}$.  What
does the function $f^{-1}$ represent?

\subsection{Combining Functions}

\question
Find $f+g$, $f-g$, $fg$ and $f/g$ and their domains:
\begin{parts}
  \part $f(x) = \sqrt{x^2-4}$ and $g(x) = \sqrt{x-2}$
  % \vspace{2cm}
  \part $f(x) = \dfrac{2}{x}$ and $g(x) = \dfrac{4}{x+4}$
  % \vspace{2cm}
\end{parts}

\question
Find $f \circ g$ and $g \circ f$, $f \circ f$, and $g \circ g$ and their domains:
\begin{parts}
  \part $f(x) = x^2$ and $g(x) = x+1$
  % \vspace{2cm}
  \part $f(x) = 6x-5$ and $g(x) = \dfrac{x}{2}$
  % \vspace{2cm}
\end{parts}

\question find $f \circ g \circ h$ where $f(x) = \sfrac{1}{x}$, $g(x) = x^2$, and $h(x) = x+3$.

\question
Find $f(x)$ and $g(x)$ where $h(x) = (f \circ g)(x)$.
\begin{parts}
  \part $h(x) = (x-3)^3$
  \part $h(x) = \dfrac{x^2}{x^2+1}$
\end{parts}

\question
A spherical balloon is being inflated at a rate which causes the radius to increase by 2 cm each second.  Express the surface area
of the baloon as a function of time.  The formula for the surface area of a sphere is $A = 4 \pi r^2$.

\question
An appliance dealer advertises a 10\% discount on all of his washing machines.  In addition, the manufacturer offers a
\$100 rebate on the purchase of a washing machine.  Let $x$ represent the sticker price of the washing machine.

\begin{parts}
  \part Supose that only the 10\% discount applies.  find a function $f$ that models the purchase price of the washer as
  a function of the sticker price $x$.

  \part Supose that only the \$100 rebate applies.  find a function $g$ that models the purchase price of the washer as
  a function of the sticker price $x$.

  \part Find $f \circ g$ and $g \circ f$.  What do they represent?  Which is the better deal?

\end{parts}

\end{questions}


\end{document}

