
\documentclass[fleqn,addpoints]{exam}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}

% \printanswers

\ifprintanswers \usepackage{2in1, lscape} \fi

\title{Math 115 Chapter One Review Problems}
\date{November 23, 2010}
\author{}

\begin{document}

\maketitle

\section{Chapter One Test}

\subsection{Overview}

The \emph{Chapter One} text will be next week.  It will include all of \emph{Chapter One} except for finding equations
of lines.  We covered this fairly thoroughly in \emph{Math 113}, and there is plenty of other material in this chapter.

\subsection{Topics}

Here are the main topics:
\begin{description}
\item[Inequalities] both with and without absolute value
\item[Equations of Circles] $(x-h)^2 + (y-k)^2 = r^2$
\item[Functions] and their domain and range
\item[Rigid Transformations] Given the graph for $f(x)$, draw the graphs of $f(x-c)$, $f(x+c)$, $f(x) + c$, $f(x) - c$,
  and $cf(x)$
\item[Combinations of Functions] $(f+g)(x)$, $(f-g)(x)$, $(fg)(x)$, $(f/g)(x)$, and $(f \circ g)(x)$
\item[Inverse Functions] determine if a function has an inverse and find $f^{-1}(x)$
\end{description}

\pagebreak

\section{Review Problems}

You don't have to hand these problems in.  But you should know how to do them all to prepare for the test.

\subsection{From the Book}

\begin{itemize}
  \item pp. 101-105 (Review Exercises): 9, 12, 14, 26-28, 44-45, 50-54, 56-57, 61-62, 65-68, 70, 75, 78-79, 81
\end{itemize}

\subsection{Other Problems}

\begin{questions}
\question
Find $f+g$, $f-g$, $fg$ and $f/g$ and their domains:
\begin{parts}
  \part $f(x) = \dfrac{2}{x+1}$ and $g(x) = \dfrac{x}{x+1}$
  \part $f(x) = \sqrt{x^2-1}$ and $g(x) = \sqrt{x+1}$
\end{parts}

\question
Find $f \circ g$ and $g \circ f$ and their domains:
\begin{parts}
  \part $f(x) = x^3 + 2$ and $g(x) = \sqrt[3]{x}$
  \part $f(x) = \dfrac{x}{x+1}$ and $g(x) = 2x-1$
\end{parts}

\question
Find $f(x)$ and $g(x)$ where $h(x) = (f \circ g)(x)$ (there is more than one answer)
\begin{parts}
  \part $h(x) = |1 - x^2|$
  \part $h(x) = \sqrt{x + \sqrt{x}}$
\end{parts}

\question

Find $f^{-1}(x)$
\begin{parts}
  \part $f(x) = 3x-1$
  \part $f(x) = \dfrac{x-2}{x+2}$
  \part $f(x) = x^2-4$, $x \geq 0$
\end{parts}

\end{questions}

\ifprintanswers

\fi

\ifprintanswers
\else
\vspace{7 cm}

{\em I'm the one that's going to have to die when it's time for me to die.  So let me live my life the way I want to.}

\vspace{.1 cm}
\hspace{1 cm} --Jimi Hendrix, {\em If 6 was 9}

\fi

\end{document}

