% no answer key
% \documentclass[letterpaper]{exam}

% answer key
\documentclass[letterpaper, landscape]{exam}
\usepackage{2in1, lscape} 
\printanswers

\usepackage{units} 
\usepackage{xfrac} 
\usepackage[fleqn]{amsmath}
\usepackage{commath}
\usepackage{cancel}
\usepackage{float}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{mathtools} 

\newcommand{\degree}{\ensuremath{^\circ}} 
\everymath{\displaystyle}

\title{Math Prep II \\ Homework One}
\author{}
\date{\today}

\begin{document}

  \maketitle

  \section{From the Book} % (fold)
  \label{sec:Homework}
  
  \begin{itemize*}
    \item pp 9-10, problems 1-6, 36-38
    \item pp 40-42, problems 1, 15-20, 25-30, 33-38, 53, 56
  \end{itemize*}

  \section{Problems}
  \begin{questions}

    \question

    In 1772, the German astronomer Johann Elert Bode found a pattern in the distances of the planets
    from the sun.  At the time, only six planets were known.  The actual relative distances of the
    planets from the sun and his pattern are:

    \begin{tabular}{lrr}
      \toprule
      Planet  & Actual Distance & Bode's Pattern \\ 
      \midrule
      Mercury & 4               & \( 0 + 4 = 4 \) \\
      Venus   & 7               & \( 3 + 4 = 7 \) \\
      Earth   & 10              & \( 6 + 4 = 10 \) \\
      Mars    & 15              & \( 12 + 4 = 16 \) \\
      ?       & ?               & ? \\
      Jupiter & 52              & \( 48 + 4 = 52 \) \\
      Saturn  & 96              & \( 96 + 4 = 100 \) \\
      \bottomrule
    \end{tabular}

    \begin{parts}
      \part What equation belongs between Bode's equations for Mars and Jupiter?

      \begin{solution}
        After the first equation, the equations all look like: 
        \[ 
          2^n \cdot 3 + 4 
        \]
        where n = \{ 0, 1, 2, \ldots \}.  

        The missing equation is: $24 + 4 = 28$

      \end{solution}
      \part What equation belongs after Bode's equation for Saturn?

      \begin{solution}
        Using the same reasoning, the missing equation is: $192 + 4 = 196$

      \end{solution}

    \end{parts}

    In 1781 William Herschel discovered Uranus, the next planet beyond Saturn.  Because its distance
    of 192 units comes remarkably close to the number predicted by the equation, astronomers came to
    the conclusion that the equation between Bode's equations for Mars and Jupiter must also mean
    something.  In fact, in 1801, the asteroid Ceres was discovered at a distance of 28 units from
    the sun.

    \question

    In the late sixteenth century, the Italian scientist Galileo discovered an interesting
    relationship between the length of a pendulum and the time of the swing:

    \begin{tabular}{cc}
      \toprule
      Length of Pendulum & Time of Swing \\
      \midrule
      1 unit             & 1 second \\
      4 units            & 2 seconds \\
      9 units            & 3 seconds \\
      16 units           & 4 seconds \\
      \bottomrule
    \end{tabular}

    \begin{parts}
    \part
    Write an equation for the length of the pendulum in terms of the time of the swing.

    \begin{solution}
    The length is always the time of the swing squared.  So the equation is: \( L = T^2 \)
    \end{solution}

    \part
    What would be length of a pendulum with a swing of 8 seconds?

    \begin{solution}
    Substituting 8 into the equation gives: 
    \begin{eqnarray*}
      L & = & T^2 \\
        & = & 8^2 \\
        & = & 64 \\
    \end{eqnarray*}
    \end{solution}

    \end{parts}

    \section{Extra Credit}

    \question

    You have 9 metal balls and a balance which allows you to compare the weight of two balls or two
    groups of balls.  You know that one of the balls is slightly heaver than the others but don't know
    which one is the heavier ball.  What is the smallest number of weighings required to identify the
    heavier ball?

    \begin{solution}
      To minimize the number of weighings, you want to eliminate as many balls as possible with each
      weighing.  This approach allows you to eliminate two thirds of the balls each time, resulting in
      two weighings.

      \begin{itemize}
        \item 
          Divide the balls into three groups of three.  Select two of the groups and put them on the
          balance.  If one of the groups is heavier, the heavy ball is in that group.  Otherwise, the
          heavy ball is in the group you didn't weigh.

        \item 
          Select two balls from the group containing the heavy ball and put them on the balance.  If
          one of them is heavier, it, obviously, is the heavy ball.  Otherwise, the third ball is the
          heavy ball. 
      \end{itemize}

      This procedure works for any number of balls that is a power of three. The number of weighings
      required is equal to the exponent:

      \begin{tabular}{rrr}
        \toprule
        Exponent & Number of Balls & Number of Weighings \\
        \midrule
        0        & $3^0 = 1$       & 0 \\
        1        & $3^1 = 3$       & 1 \\
        2        & $3^2 = 9$       & 2 \\
        3        & $3^3 = 27$      & 3 \\
        4        & $3^4 = 81$      & 4 \\
        \vdots   & \vdots          & \vdots \\
        \bottomrule
      \end{tabular}

    \end{solution}

    \question

    You throw away your balance and buy a scale instead.  But you are now faced with a different problem.  

    You have 10 stacks of coins, each of which contains 10 coins.  You know that one stack is full of
    counterfeit coins while all the other stacks contain real coins.  You also know that a real coin
    weighs 10 ounces while a counterfeit coin weighs 11 ounces.  What is the smallest number of
    weighings required to identify the stack of counterfeit coins?

    \begin{solution}

      The best approach only requires one weighing.  The key is to notice that you can select
      different numbers of coins from each stack.  So what you do is select 1 coin from the first
      stack, 2 coins from the second stack, 3 coins from the third stack, etc. and weigh them all at
      once.

      If all of the coins were real coins, you would expect the total to be: 
      \[ 
        10 \del{ 1 + 2 + 3 + 4 + 5 + 6 + 7 + 8 + 9 + 10 }  = 10 \cdot 55 = 550 
      \]

      The actual total will be more than the expected total by the stack number.  

      For example, suppose stack 7 contains the counterfeit coins.  The seven coins you included from
      stack 7 all weigh one ounce more than they should, so the total in this case will be: 
      \[ 
        550 + 7 = 662 
      \]

      To get the number of the stack containing the counterfeit coins, you just need to subtract 550
      from whatever it says on the scale.

    \end{solution}

  \end{questions}

\end{document}


