% % no answer key
% \documentclass[letterpaper]{exam}

% answer key
\documentclass[letterpaper, landscape]{exam}
\usepackage{2in1, lscape} 
\printanswers

\usepackage{units} 
\usepackage{xfrac} 
\usepackage[fleqn]{amsmath}
\usepackage{cancel}
\usepackage{float}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{parskip}

\everymath{\displaystyle}


\title{Statistics \\ Homework Seven}
\date{\today}
\author{}

\begin{document}

  \maketitle

  \section{Homework}
    \begin{itemize*}
      \item read Chapter 7 
      \item take a look at the ``Check Your Skills'' exercises
      \item exercises: TO DO
    \end{itemize*}

  \ifprintanswers
    \begin{description}

      \item[25] The population is adults 18 and over and the sample is the 1027
        people actually interviewed.

      \item[27] If you number the students from 01 to to 40 and use Table B
        starting at line 117, you get students: 38, 16, 32, 18, 37, 06, 23,
        19, 03 and 25.

        This is students: Washburn, Garcia, Rodriguez, Helling, Wallace,
        Cabrera, Morgan, Husain, Batista, Nguyen

      \item[28] The first five codes would be: codes 288, 229, 131, 303, and 007.

      \item[32] 
        \begin{parts}
          \part There will typically be around four 0's in each row but some
            rows will have more and some less.  The mean number of zeros per row
            will be 4.

          \part Since there are 100 different pairs of digits, any particular
          pair has a 1/100 chance of appearing, including the pair 00.

          \part This pattern will come up once in a while but not very
            frequently.
        \end{parts}

      \item[33]
        \begin{parts}
          \part The population is all the residential households.

          \part 1169 out of 2000 or 58\% didn't respond.

          \part People are likely to telescope the events, thinking that they've
            seen more movies in the last 12 months than they actually have.
        \end{parts}

      \item[34] The online poll only included people who were watching Lou
        Dobb's show.  Lou's audience is primarily composed of people who think
        issuing driver's licenses to illegal immigrants is a bad idea, so they
        don't represent the population as a whole.

      \item[35] 40,927 out of 45,956 or 89\% didn't respond.

      \item[36] 
        \begin{parts}
          \part If you assign each driver a number between 0001 and 5024 and use
          line 104 of Table B, the first 5 drivers are: 1388, 0746, 0227, 4001
          and 1858.

          \part People generally don't want to admit they ran a red light, so the
          actual number is probably higher.
        \end{parts}

      \item[38] Everybody has a 1 in 10 chance of being interviewed.  This is a
        stratified sample since the students are divided into two groups
        according to age.

      \item[41] 
        \begin{parts}
          \part The only thing you need Table B for in this example is the
          starting number.  The first number less than 200 is 193, so you would
          start with name 193, and then select names 393, 593, 793, and 993.

          \part There could be some repeating pattern in the data which shows up
          when you use a systematic sample.  For example, suppose you are
          sampling houses and there are 5 houses per block.  If you start with
          house 1 and select every 10th house, you will always be selecting a
          corner house and corner houses might not be representative of all the
          houses.
        \end{parts}

      \item[42]
        \begin{parts}
          \part Houses that don't have land-line telephones or have unlisted
          numbers are omitted.

          \part By selecting random numbers, you can include the people with
          unlisted numbers.  You'll still miss the people without phones, of
          course.
        \end{parts}

      \item[43]
        \begin{parts}
          \part As described in question 42, random digit dialing means that the
          computer dials digits at random instead of going through a phone
          directory.

          \part Selecting the adult with the most recent birthday takes a SRS of
          the adults in the household.  If you don't do this, your sample might
          be biased because you mostly get the people who are most likely to answer
          the phone and this group might not be representative.
        \end{parts}

      \item[47]
        \begin{parts}
          \part The sampling error for Hispanics is higher because there are
          fewer of them in the sample.  The smaller your sample, the less likely
          it is that the sample is representative of the population.

          \part There are even fewer Cubans than Hispanics, so the sample is so
          small that you can't draw any reliable conclusions about the
          population from the sample.
        \end{parts}

  \end{description}

  \else
    \vspace{10 cm}
    \begin{quote}
      \begin{em}
      \end{em}
    \end{quote}
    \hspace{1 cm} --Howard Zinn
  \fi

\end{document}

