\documentclass[landscape]{exam}

\usepackage{2in1, lscape} 
\usepackage{units} 
\usepackage[fleqn]{amsmath}
\usepackage{float}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{parskip}

\printanswers

\everymath{\displaystyle}

\printanswers

\title{Statistics \\ Week Sixteen}
\date{\today}
\author{}

\begin{document}

  \maketitle
  \tableofcontents

  \section{Inference Cautions}
  You can't use inference when your sample is not an SRS or the outcome of a
  randomized controlled experiment.

  If your population doesn't have a Normal distribution, it's usually fine as
  long as you take a large enough sample. Because of the Central Limit Theorem,
  the values of $\bar{x}$ do have a Normal distribution if your sample is
  reasonably large.

  \section{T-test}
  Gosset was an employee of the Guinness Brewery published under name
  ``Student'' so competitors might underestimate the value of the work.

  If you don't have the standard deviation of the population, you can use the
  standard deviation of the sample. You get a t-test instead of a z-test.

  The probability density graph for a t-test is slightly different from a Normal
  graph, with the differences diminishing as the sample size increases.

  \section{Confidence Intervals}
  A {\em Confidence Interval} is in the form $x \pm y$.

  Associated with a confidence interval is:
  \begin{description}
    \item[Confidence Level] is how confident you are that the interval
      captures the actual value you are looking for. Common values are 68\%,
      95\% and 99.7\%.

    \item[Margin of Error] the width of the interval
  \end{description}

  With a larger margin of error, you can be more confident that the interval
  includes the actual value of the parameter.

  \section{Significance}
  ``Statistically significant'' means that you had a low P-value. You can be
  confident that the result didn't happen by chance.

  Statistically significant is not the same as practically significant. You may be
  very confident in a difference of 1\% (statistically significant) but not care
  because 1\% is such a small percentage (practical significant).

  If the actual effect is small, you need a large sample size to get a
  statistically significant result. This is because a large sample size results
  in a small standard deviation, which allows you to pin down the effect you are
  looking for.

  \section{Calculating Required Sample Size}

  Calculate how many samples you need for a desired margin of error, $m$:

  \begin{align*}
    m        & = \frac{z^* \sigma}{\sqrt{n}} \\
    \sqrt{n} & = \frac{z^* \sigma}{m} \\
    n        & = \left( \frac{z^* \sigma}{m} \right)^2 \\
  \end{align*}

  For example, suppose 
  \begin{align*}
    m      & = 4 \\
    \sigma & = 20 \\
  \end{align*}
  
  And you want to have 95\% confidence.  95\% is two standard deviations, so
  $z^* = 2$. 

  \begin{align*}
    n & = \left( \frac{z^* \sigma}{m} \right)^2 \\
      & = \left( \frac{2 \cdot 20}{4} \right)^2 \\
      & = 100 \\
  \end{align*}

  \section{Power}
  The power of a test is the chance that it will correctly reject the null
  hypothesis when the null hypothesis is false.

  exercise 48:

  \begin{enumerate}[(a)]
    \item 
      \begin{align*}
        z       & = 1.6449 \\
        \bar{x} & = \frac{1.6449}{3.162} \\
                & = 0.5202 \\
      \end{align*}

      Any $\bar{x}$ greater than 0.52 will reject the hypothesis.

    \item The probability of observing these values when the mean is 0.8 is:
      \begin{align*}
        z        & = \frac{0.5202 - 0.8}{1/\sqrt{10}} \\
                 & \approx -0.8848 \\
        pnorm(z) & \approx 0.8819 \\
      \end{align*}

      The probability of a value at least this large when the mean is 0.8 is
      approximately 0.8819.

      This is the same answer my computer gave, but slightly different from the
      book answer. 

  \end{enumerate}

  % The general formula for power is something like:
  % \begin{align*}
  %   z_{effect} & = \frac{effect}{\sigma/\sqrt{n}} \\
  %   z_{\alpha} & = qnorm(0.05) \\
  %   \\
  %   P(\text{reject } H_0 | H_0 \text{ false}) 
  %       & = pnorm(z_{\alpha} + z_{effect}) \\
  % \end{align*}

  You can get high power (high confidence in the outcome) by:
  \begin{itemize*}
    \item larger sample size
    \item null hypothesis very wrong so there is a large difference between the
      sample and the actual parameter (higher $z_{effect}$)
    \item lower desired significance level (higher $z_{\alpha}$
  \end{itemize*}

  \section{Type I and II Errors}
  \begin{description*}
    \item[Type I] reject $H_0$ when it is actually true.
    \item[Type II] accept $H_0$ when it is actually false.
  \end{description*}

  \begin{tabular}[H]{lll}
     hypothesis & reject          & accept \\
     true       & Type I Error    & correct outcome \\
     false      & correct outcome & Type II error \\
  \end{tabular}

  \section{Check Your Skills}
  \begin{description}
    \item[15.1] c---not a random sample

    \item[15.2] c---not a random sample
      \begin{enumerate}[(a)]
        \item 
          For a 95\% confidence interval, we need to within two standard
          deviations of the mean.

          \begin{align*}
            \sigma_{\bar{x}} & = \frac{1.83}{\sqrt{880}} \\
                             & \approx 0.06169 \\
            \\
            z^* &\approx 2 \\
            \\
            \mu & = 1.92 \pm 2 \cdot 0.06169 \\
                & \approx 1.92 \pm 0.1209 \\
                & \approx (1.80, 2.04) \\
          \end{align*}

        \item The distribution of the sample means is still normal since we have
          a large sample (Central Limit Theorem).

        \item 
          \begin{itemize*}
            \item it only includes people with phones
            \item it only includes people who answer their phone
          \end{itemize*}
      \end{enumerate}

    \item[15.3] Different times of day get different types of people.

    \item[15.4] 
      \begin{enumerate}[(a)]
        \item

          \begin{table}[ht]
          \centering
          \begin{tabular}{rrrrrrr}
            \toprule
              & p    & z    & s.z  & margin.of.error & min   & max \\
            \midrule
            1 & 0.90 & 1.64 & 0.29 & 0.48            & 26.32 & 27.28 \\
            2 & 0.95 & 1.96 & 0.29 & 0.57            & 26.23 & 27.37 \\
            3 & 0.99 & 2.58 & 0.29 & 0.76            & 26.04 & 27.56 \\
            \bottomrule
          \end{tabular}
          \end{table}
      \end{enumerate}

  \end{description}
\end{document}

