% no answer key
% \documentclass[letterpaper]{exam}

% answer key
\documentclass[letterpaper, landscape]{exam}
\usepackage{2in1, lscape} 
\printanswers{}

\usepackage{units} 
\usepackage{xfrac} 
\usepackage[fleqn]{amsmath}
\usepackage{commath}
\usepackage{cancel}
\usepackage{float}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{mathtools} 
\usepackage{parskip} 

\newcommand{\degree}{\ensuremath{^\circ}} 
\everymath{\displaystyle}

\title{Algebra \\ Homework 7 \\ Sections 3.1 and 3.2 }
\author{}
\date{\today}


\begin{document}

  \maketitle

  \section{Homework}

  \begin{itemize*}
      \item Section 3.1: 11--20, 26--30, 36--40, 36--40, 50--55, 61--68
      \item Section 3.2: 5--10, 21--25, 50--55, 65--70, 81--84, 92
  \end{itemize*}

  \section{Extra Credit}

  A silver prospector was unable to pay his March rent in advance.  He owned a bar of pure silver,
  31 inches long, so he made the following arrangement with his landlady.  He would cut the bar, he
  said, into smaller pieces.  On the first day of March he would give the lady an inch of the bar,
  and on each succeeding day he would add another inch to the amount of silver.  She would keep this
  silver as security.  At the end of the month, when the prospector expected to be able to pay his
  rent in full, she would return the pieces to him.

  March has 31 days, so one way to cut the bar would be to cut it into 31 sections, each an inch
  long.  But since it required considerable labor to cut the bar, the prospector wished to carry out
  his arrangement with the fewest possible number of pieces.  For example, he might give the lady an
  inch on the first day, another inch the second day, then on the third day he could take back the
  two pieces and give her a solid 3-inch section.

  Assuming the portions of the bar are traded back and forth in this fashion, what is the smallest
  number of pieces into which the prospector needs to cut his silver bar?

  (Martin Gardner)

  \begin{solution}

      The most efficient approach is that each time he runs out of pieces he cuts off a new piece
      exactly as long as he needs to meet that day's rent.  

      \begin{itemize}
          \item On the first day, he has zero pieces, so he cuts off a piece of size 1.  

          \item On the second day, he needs a piece of size 2, so he cuts off a piece of size 2.

          \item On the third day, he gives the landlady both pieces.

          \item On the fourth day, he needs another piece, so he cuts off a piece of size 4.

          \item On days 5--7, he uses the pieces he already has.

          \item On the eighth day, he needs another piece, so he cuts off a piece of size 8.

          \item On days 9--15, he uses the pieces he already has.

          \item On the sixteenth day, he needs another piece. The rest of the bar is size 16, so he
            uses that.

          \item On days 17--31, he uses the pieces he already has.
      \end{itemize}

      He cuts the bar into pieces of size 1, 2, 4, 8, and 16, and four cuts are required.

  \end{solution}
  \ifprintanswers{}

      \section{Section 3.1} 

      \begin{description}

          \item[11] $10x - 3$

          \item[12] $14x + 3$

          \item[13] $-11t + 5$

          \item[14] $-10t + 8$

          \item[15] $-x^2 + 2x - 2$

          \item[16] $-x^2 + x - 6$

          \item[17] $17a^2b^2 - 5ab$

          \item[18] $-5a^2b^2 - 7ab$

          \item[19] $-9x + 7$

          \item[20] $-2x^2 -2x - 5$

          \item[26] $-2x^2 - 2x - 2$

          \item[27] $-6a^2 + 12a + 14$

          \item[28] $6a^2 + 12a - 14$

          \item[29] $3x^3 + x^2 + 13x - 11$

          \item[30] $3x^3 - x^2 - x - 6$

          \item[36] $-5x^2 -3x + 16$

          \item[37] $-3x^3 + 5x^2 - 2x + 9$

          \item[38] $-x^3 + 4x^2 + x - 5$

          \item[39] $5x^2 - 4x + 11$

          \item[40] $-x^3 - 2x^2 + 7x - 2$

          \item[50] $3x + 7$

          \item[51] $-9x^2 -12x + 4$

          \item[52] $-10x^2 + 2x + 14$

          \item[53] $10x^2 + 13x - 18$ 

          \item[54] $-3x^2 - 2x + 10$

          \item[55] $-n^2 -4n -4$

          \item[61] $-7n^2 + n + 6$

          \item[62] $-11n^2 - n + 4$

          \item[63] $t^2 - 4t + 8$

          \item[64] $-4n^2 + 3n - 1$

          \item[65] $4n^2 - n - 12$

          \item[66] $6$

          \item[67] $-4x - 2y$

          \item[68] $7xy - 2x + 3y$

      \end{description}

      \section{Section 3.2}

      \begin{description}
          \item[5] $4a^3b^4$

          \item[6] $24a^3b^5$

          \item[7] $-3x^3y^2z^6$

          \item[8] $2x^3y^5z^3$

          \item[9] $-30xy^4$

          \item[10] $-28x^5y$

          \item[21] $30x^6$

          \item[22] $12x^6$

          \item[23] $-18x^9$

          \item[24] $-84x^6$

          \item[25] $-3x^6y^6$

          \item[50] $16a^8b^{16}$

          \item[51] $-16a^4b^4$

          \item[52] $-81a^4b^4$

          \item[53] $-x^6y^{12}z^{18}$

          \item[54] $-x^8y^{16}z^{24}$

          \item[55] $-125a^6b^6c^3$

          \item[65] $-18xyz^4$
          
          \item[66] $-32x^2y^4z^5$

          \item[67] $-a^2b^3c^2$

          \item[68] $-a^2b$

          \item[69] $9$

          \item[70] $-8y$

          \item[81] $a^{5n+1}$

          \item[82] $x^{3n}$

          \item[83] $-10x^{2n}$

          \item[84] $-12x^{3n}$ 

          \item[92]
              The sides come in pairs: 
              \begin{itemize*}
                \item $2x^2$
                \item $5x$
                \item $10x$
              \end{itemize*}

              \begin{itemize}
                \item Area: $2(2x^2) + 2(5x) + 2(10x) = \boxed{ 4x^2 + 30x }$
                \item Volume: $x \cdot 2x \cdot 5 = \boxed{ 10x^2 }$
              \end{itemize}

      \end{description}

  \fi
  \ifprintanswers{}
  \else
    \vspace{3 cm}
    \begin{quote}
      \begin{em}
        On two occasions I have been asked [by members of Parliament], ``Pray, Mr.\ Babbage, if you
        put into the machine wrong figures, will the right answers come out?'' I am not able
        rightly to apprehend the kind of confusion of ideas that could provoke such a question.
      \end{em}
    \end{quote}
    \hspace{2 cm}--Charles Babbage
  \fi

\end{document}

