\documentclass[letterpaper, landscape]{exam}
\usepackage{2in1, lscape} 
\printanswers

\usepackage{units} 
\usepackage{xfrac} 
\usepackage[fleqn]{amsmath}
\usepackage{commath}
\usepackage{cancel}
\usepackage{float}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage{cancel}
\usepackage{polynom}
\usepackage{caption}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{mathtools} 

\newcommand{\degree}{\ensuremath{^\circ}} 
\everymath{\displaystyle}


\title{Math Prep II \\ Week One Notes}
\author{}
\date{\today}

\begin{document}

  \maketitle

  \section{Sets}

  \begin{itemize}
    \item A {\em set} is an unordered collection of elements.  

    \item Two sets are equal if they contain the same elements.

    \item There are several ways to define a set.  These all define the set of positive odd integers less than 20:
      \begin{align*}
        A & = \cbr{ 1, 3, 5, 7, 9, 11, 13, 15, 17, 19 } \\
        A & = \cbr{ 19, 3, 11, 9, 15, 1, 13, 7, 17, 5 } \\
        A & = \cbr{ 1, 3, 5, \ldots, 17, 19 } \\
        A & = \cbr{ x | \text{ x is odd and between 1 and 20} } \\
      \end{align*}

    \item One set is a {\em subset} of another set if all of the elements in the first set are
      contained in the second set.

      \begin{align*}
        \cbr{ 3, 5 } & \subseteq \cbr{ 1, 3, 5, 7 } \\
        \cbr{ 3, 5 } & \subseteq \cbr{ 3, 5 } \\
        \emptyset    & \subseteq \cbr{ 3, 5 } \\
      \end{align*}
  \end{itemize}

  \section{Types of Real Numbers}

  For most of this course we will be using {\em real numbers}.  There are several types of real numbers:

  \begin{itemize}
    \item {\em Integers} are the numbers \( \{ \ldots, -3, -2, -1, 0, 1, 2, 3, \ldots \} \)
    \item {\em Rational Numbers} are the numbers which can be expressed as a ratio of two integers:
      1/2, 3/4, 2, 3.5, etc.
    \item {\em Irrational Numbers} are numbers which can't be expressed as a ratio of two integers:
      \( \sqrt{2} \), \( \pi \), \( e \), \( \sqrt{17}\), etc.
  \end{itemize}

  \section{Absolute Value}

  The absolute value of a number is the number without its sign.

  \begin{align*}
    \envert{17}  & = 17 \\
    \envert{-17} & = 17 \\
    \envert{0}   & = 0 \\
  \end{align*}

  \section{Number Properties}

  The text book describes several properties of numbers, most of which are fairly obvious (for
  example: if \( a = b \) then \( b = a \)).  You should read over the chapter to make sure the rest
  of them are also familiar to you.  

  Here are a few of the properties which are slightly less obvious and quite useful:

  \begin{align*}
    a (b + c)   & = ab + ac \\
    a + (b + c) & = (a + b) + c \\
    (ab)c       & = a(bc) \\
  \end{align*}

  \section{Algebraic Expressions}

  An algebraic expression usually contains several {\em term}.  A {\em term} is a number and some
  variables multiplied together.  Here are some terms: \( x^2 \), $ 3xy $,  $ 47 $, $ 12x^3y^2z^4 $.


  Because of the distributive property, terms with exactly the same variables (including exponents)
  are called {\em like} or {\em similar} terms and may be added together:

  \begin{align*}
    3x^2 + 7x^2 & = x^2(3 + 7) \\
                & = 10x^2  \\
  \end{align*}

  To evaluate an algebraic expression for a particular value for variable, you substitute the value
  in for the variable everywhere the variable appears in the equation.  Usually it's easiest to
  simplify the equation as much as you can before you do this.

\end{document}


